---
title: "machinelearn 2"
---
Machine learning Assignment from Coursera
output: html_document

## Introduction

This is a study of how measures from sensors in wearable devices can predict activity. 

In this case the question is which predictors will best predict 5 weight lifting situations indicated by letters A through E.

The prediction is concerned with how well a model predicts these situations from a test set which is provided. 

The model will be built using the caret function "train" using random forests. The model will be based on training data which is 60% of the training set provided.

## Exploratory Analysis:

There are 2 data files available from the course assignment page, a training set and a testing set.

These sets were examined in R and Excel. Variables with no values or NA were removed thereby reducing the variables to 52 in each set. These were then read into R for further processing.A factor variable and numeric variable missmatch in training and pmltesting was reconciled

Read in data files from Cousera assignment page after cleaning:
```{r}
pmltraining2<-read.csv("pml-training2.csv")
pmltesting2<-read.csv("pml-testing2.csv")
#show the dimensions of the data frames
dim(pmltraining2)
dim(pmltesting2)
```

The pmltraining set was divided further into training and testing sets
using the caret function createDataPartition. These two data frames will be used for cross validation and measurement of out of sample error rate.

```{r}
library(caret)
set.seed(123)
inTrain=createDataPartition(pmltraining2$classe,p=.6[[1]],list=F)
training=pmltraining2[inTrain,]
testing=pmltraining2[-inTrain,]
dim(training)
dim(testing)
```

The outcome variable in the pmltraining2 data is called "classe". In the pmltesting2 data there is a varible called "problem_id". The other variable names are the same.

## Developing the machine learning algorithm:

To develop the algorithm I used the train function from the carat package and random forests. Using all the variables, I applied this to the training data.

Train does a Cross validation as part of the process so this will not have to done separately.

trainControl  was also set into the model.

```{r}
trControlA<-trainControl(method="cv",number=4)
modelfit<-train(classe~.,data=training, method="rf",prox=TRUE, trControl=trControlA)
```
  

## Results:

```{r}
modelfit
```



```{r}
modelfit$finalModel
```

This model did very well with an OOB error rate of 0.34%


## Predicting on the testing set:

```{r}
predtest<-predict(modelfit,newdata=testing)
confusmat<-confusionMatrix(testing$classe,predtest)
confusmat
```


##Predicting on the pmltesting2 set:

predictpmltest<-predict(modelfit, newdata=pmltesting2)




